\large{\bf Large effects and the infinitesimal model }

### 2023-07-07 writeup.tex
.
Clean up.

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
% In fact, many traits claimed to be genetically determined
% in fact only appear heritable through non-genetic means,
% e.g., systematic inequalities in resource distribution partitioned by class or race
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should \emph{not} always be the default for understanding biological variation,
especially for human traits.

@@

\comment{Proposal: change $1 + \mu$ to $\mu$ (or, something else, but a single letter).}

\comment{Copied from \texttt{eugene\_notes.tex}:}

Let $X$ be a population of size $N$, recorded as the empirical distribution on phenotypes
\begin{align*}
    X = \frac{1}{N} \sum_{i=1}^N \delta_{x_i} .
\end{align*}
So, $X$ is a point measure of total mass 1 on $\R$.
This evolves as a Moran process under the infinitesimal model
with death rates depending on state $x$.
Concretely, an individual whose phenotype is $x$
dies at rate $1 + \mu(x)$
and is replaced by a new individual of phenotype
\begin{align*}
  \frac{y + z}{2} + \xi ,
\end{align*}
where $y$ and $z$ are independent draws from $X$
and $\xi$ is an independent draw from the transmission noise distribution.
Let $p_\xi$ denote the distribution of $\xi$,
and $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
This has generator
\begin{align*}
    \mL \Phi(X)
    &=
    \int \int \int \int
    \left( \Phi\left( X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x \right) - \Phi(X) \right)
    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
    &=
    \int \int
    \left( \Phi\left(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x\right) - \Phi(X) \right)
    (1 + \mu(x)) X(dx) \nu(X)(du) .
\end{align*}

For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X} := \int \phi(x) X(dx)$
we get the first moment
\begin{align*}
    \mL \ip{\phi}{X}
    &=
    \frac{1}{N} \int \int \int \int
    \left( \phi \left( \frac{y+z}{2} + \xi \right)
         - \phi(x) \right)
    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
    &=
    \frac{1}{N} \ip{\phi}{\nu(X)}\ip{1+\mu}{X}
    - \frac{1}{N} \ip{\phi}{(1+\mu) X} .
\end{align*}
Note that for $Y$ another finite measure on $\R$,
the Fr\'echet derivative \comment{(is that the right name?)} of $X$ in the direction of $Y$
satisfies $\ip{\phi}{\partial_Y X} = \lim_{\epsilon \to 0} \epsilon^{-1} (\ip{\phi}{X + \epsilon Y} - \ip{\phi}{X}) = \ip{\phi}{Y}$,
so that this is a first derivative in the direction of the measure
$$ \ip{1+\mu}{X} \nu(X) - (1+\mu) X . $$

More generally, if $\Phi(X) = \prod_i \ip{\phi_i}{X}$ then
\begin{align*}
    \mL \Phi(X)
    &=
    \frac{1}{N} \sum_i \left( 
        \ip{1+\mu}{X} \ip{\phi_i}{\nu(X)}  - \ip{\phi_i}{(1+\mu)X}
    \right)
    \prod_{j \neq i} \ip{\phi_j}{X} 
    + O(1/N^2) .
\end{align*}
This implies that after rescaling time by $N$
and suitable other caveats,
the limiting process is a deterministic flow in the direction of 
$\nu(X) \ip{1+\mu}{X} - (1+\mu)X$.

\paragraph{Stable measures}
If $Z$ is a fixed point of this flow,
then $\nu(Z)\ip{1+\mu}{Z} = (1+\mu)Z$.
In other words, any measure $Z$ that is a fixed point has the property that
(the average of two independent draws plus transmission noise)
is equal in distribution to
(a $1+\mu$-weighted draw).
Concretely, with $A$, $B$, and $C$ independent draws from $Z$,
\begin{align} \label{eqn:ABC_condition}
    \E\left[ f\left( \frac{A+B}{2} + \xi \right) \right]
    &=
    \frac{\E\left[ f(C) (1+\mu(C)) \right]}{\E[1+\mu(C)]} .
\end{align}

\comment{Copied from \texttt{MeanFlow.tex}:}

Assume that $Z$ is a finite measure, and without loss of generality normalized to have total mass one.

Throughout, I'll be using the characteristic function of $Z$, which I'll denote by 
\[
	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right],
\]
and the characteristic function of $\xi$, $\Phi(\theta)$.

\section{No Selection}

First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{CHARACTERISTIC}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta),
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta).
\]
In particular, we can characterize all possible fixed points by considering the iterates of an arbitrary characteristic function, $\varphi(\theta)$, for  any distribution on $\mathbb{R}$.

Iterating $F$ $n$ times, we get
\[
	F^{(n)}(\varphi)(\theta)
	=  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
\]
so that
\[
	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
\]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \ln{\varphi}(\theta)$ and $\Psi(\theta) = \ln{\Phi(\theta)}$ exist in some neighbourhood of 0 as well.  From the above, we have
\[
	\ln{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]

\subsection{Noise Terms}

We now analyse the component arising from the noise $\xi$.  First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$
and variance $\eta^{2} = \Psi''(0)$.   Taylor's theorem then tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) = m\theta + \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Then,
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(m \frac{\theta}{2^{k}} - \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&= (n-1) m - \eta^{2} \theta^{2} (1 - 2^{-n+1}) 
		+ \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
 \end{align*}

Immediately, we see that for the limit to exist, we must have $m = 0$; if we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. We note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
\[
	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
We then have
\begin{multline*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	= 
	\begin{cases}
		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
		 \begin{multlined}
		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
		\right)    
		\end{multlined}
		& \text{if $\alpha = 1$.}
	\end{cases}
\end{multline*}
This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha \ge 1$, in which case the limit is 
\[
	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\]
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

\subsection{Reproductive Terms}

We now consider the terms $2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$, which we call the reproductive terms.  Proceeding as below, we consider the limit
%We note that this does not exclude the possibility of distributions without a mean; % \eg if $\eta =0$, then \eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.%  One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$. 
%Now, to exhaust all possibilities, observe that by iterating, we we also have that 
%\[
%	\psi(\theta) = 2^{n}\psi\left(\frac{\theta}{2^{n}}\right) 
%\]
%\ie we must have 
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm characteristic function, it must be uniformly continuous and vanish at 0, and thus $\alpha(\theta)$ as well.  We also have that for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

Now, from the above, 
\[
	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
\]
as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
\ie if $\gamma > 0$, then $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ is the characteristic function of a Cauchy distribution and $\phi(\theta)$ is the characteristic function of the sum of Cauchy and Gaussian random variables, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, so the limiting random variable is the Gaussian $N(m,\eta^{2})$.

Finally, we remark that if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist.

\subsection{Other Solutions?}

We would naturally like to know if these two exhaust the possible solutions.  To this end, we recall Bochner's theorem, which tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
\begin{enumerate}[(i)]
\item $\phi(\theta)$ is uniformly continuous,
\item $\phi(0) = 1$,
\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries
\[
	a_{ij} = \phi(\theta_{i}-\theta_{j})
\]
is positive definite.
\end{enumerate}
Given $\alpha(\theta)$ as above, the corresponding function $\phi(\theta) = e^{\alpha(\theta)-\eta^{2}\theta^{2}}$ 
or $\phi(\theta) = e^{\alpha(\theta) - \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right)}$ satisfies the first three criteria.  We will now seek to identify which $\alpha(\theta)$ result in positive definite $\phi(\theta)$.  When $\alpha(\theta) = im \theta -\gamma|\theta|$ for $\gamma > 0$, this is immediate, as $e^{\alpha(\theta)}$ is then the characteristic function of a Cauchy random variable.

We can also easily exclude the case $\gamma < 0$ using a simple consequence of positive definiteness in the case when $n=2$ (name criterion): if $\phi(\theta)$ is positive definite, then $|\phi(\theta)| \leq 1$.  For us, this requires that 
\[
	\Re \alpha(\theta) \leq  \eta^{2} \theta^{2},\; resp. \; < \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}
\]
which is violated for small values of $\theta$ when $\gamma < 0$ (recall $\alpha > 1$).

\noindent\emph{Question:} how can we use positive definiteness to exclude $\alpha(\theta)$ with $\Re \alpha(\theta) \leq  \eta^{2} \theta^{2}$ so that $\alpha'(0+)$ does not exist?

\begin{rem}
Polya has shown that if $\phi(\theta)$ is convex and satisfies the first three criteria of Bochner's theorem, then $\phi(\theta)$ is a characteristic function.  In our case, if $\alpha(\theta)$ is convex, then $e^{\alpha(\theta)}$ is as well; a proof-by-picture, however, shows that the only choice of $\alpha(\theta)$ that is convex (or concave, for that matter) for $\theta > 0$ is a linear function, so Polya's criterion doesn't give us any new characteristic functions.
\end{rem}

@@

@@

\bigskip

\noindent\emph{Question:} other stable laws?

>>>>

\DeclareMathOperator*{\Log}{Log}
\DeclareMathOperator*{\sgn}{sgn}

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should \emph{not} always be the default for understanding biological variation,
especially for human traits.

@@

As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.  The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution (\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$ is approximately $\frac{1}{2\pi\sigma^{2}} e^{-\frac{(z-m)^{2}}{2\sigma^{2}}}$ for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$.  Leaving aside -- as the Gaussian-population approximation frequently does -- the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if) deterministic mutation selection-balance is possible in an non-Gaussian model.

Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$ and is replaced by a new individual of phenotype
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution and $R$ is an independent draw from the transmission noise distribution (independent of $Z^{1}$ and $Z^{2}$).  Mutation selection balance demands that \comment{I'm struggling a bit to find a clear, short, and simple argument to jump directly to this point} that $\frac{Z^{1} +Z^{2}}{2} + R$ be equal in distribution to the selection-biased choice of $Z$, \ie for all bounded and continuous functions $f$,
\begin{equation} \label{eqn:MSB_condition}
    \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
\end{equation}
where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.

Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
\begin{align} \label{eqn:MSB_CF}
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
	\frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	&= \E\left[ e^{i\frac{theta}{2} Z^{1}}e^{i\frac{theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{\i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)	^{2} \Phi(\theta),
\end{align*}
where
\begin{equation}\label{eq:phi}
	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right]
\end{equation}
and
\begin{equation}\label{eq:Phi}
	\phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right]
\end{equation}
are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.

\subsection{The Neutral Case}

First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta),
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$:
\[
	\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta).
\]
Iterating $F$ $n$ times, we get
\[
	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
\]
so that
\[
	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
\]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We will briefly explore what limiting values are possible -- and which limits are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}. For the sake of simplicity, we will limit ourselves to the caae when both have well-posed limits independent of the other.

\subsubsection{Reproductive Terms}

First consider the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm characteristic function, it must be continuous in some neighbourhood of, and vanish at 0, and thus $\alpha(\theta)$ as well.  Furthermore, for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

Now, from the above, 
\[
	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
\]
as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
Now, if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
\[
	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1
\]
so $\gamma \geq 0$. If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

\paragraph{Lamarkian inheritance redux.} 
Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait \ie Lamarkian blending inheritance.  The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.  The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieved a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.

\paragraph{Other solutions?}
We would naturally like to know if these two exhaust the possible solutions.  Above, we exhausted all solutions for which $\alpha'(0+)$ exists.  However, if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist. This does not, however, in and of itself exclude $e^{\alpha(\theta)}$ from being a characteristic function: Bochner's theorem tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
\begin{enumerate}[(i)]
\item $\phi(\theta)$ is continuous,
\item $\phi(0) = 1$,
\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
\end{enumerate}
One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.

\subsubsection{Noise Terms}

We now turn our attention to the noise terms.  A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.  Taylor's theorem then tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

It is unclear what more can be said in general. If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.  We recall that if $\Psi(\theta)$ corresponds to a stable law(see \eg \cite{breiman1992probability}), then
\[
	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	&= \begin{cases}
		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
		 c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n(n-1)}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
	\end{cases}
\end{align*}
which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
\[
	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\]
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

More generally, one will get convergence if the noise is a mixture of components with finite variance and stable laws as above, whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms -- we remark that in the presence of noise, a pure Cauchy distributed solution is \emph{not} possible in the neutral case.

@@

Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.  We will content ourselves with presenting two examples in the case of stabilizing selection, one already familiar to enthusiasts of quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.

@@
